---
title: "Results Section"
author: "Dave Dowding"
date: "21/04/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# CHAPTER 4 - RESULTS
## 4.1 Introduction
The results of the methods described in the previous chapter are set out under the same themes that were used in the methods section namely completeness, thematic accuracy, temporal quality, positional accuracy and contributors. The results for each section are displayed and then discussed. The chapter finishes with a discussion about what together the results mean for OSM as a data source for visually impaired people, this is done under the title of ‘usability’.

## 4.2 Completeness

*OSM data compared with the Ordnance Survey dataset*

```{r}
library(plyr)
library(knitr)
os_results_df=read.csv("../dissertation/export_all_points.csv", header=TRUE, stringsAsFactors = FALSE)

results=ddply(os_results_df,~City+Type,summarise,'Total'=sum(!is.na(City)),'< 20m'=sum(Points.Distance<20),'% <20m'=round((sum(Points.Distance<20))/(sum(!is.na(City)))*100,digits=2),'No. < 50m'=sum(Points.Distance<20),'% < 50m'=round((sum(Points.Distance<50))/(sum(!is.na(City)))*100,digits=2))
kable(results,caption="Table x - showing ....")
#os_results_df
```

