---
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# CHAPTER 4 - RESULTS
## 4.1 Introduction
The results of the methods described in the previous chapter are set out under the same themes that were used in the methods section namely completeness, thematic accuracy, temporal quality, positional accuracy and contributors. The results for each section are displayed and then discussed. The chapter finishes with a discussion about what together the results mean for OSM as a data source for visually impaired people, this is done under the title of ‘usability’.

## 4.2 Completeness

*OSM data compared with the Ordnance Survey dataset*

```{r}
library(plyr)
library(knitr)
os_results=read.csv("../data/results/os_osm.csv", header=TRUE, stringsAsFactors = FALSE)

results=ddply(os_results,~City+Type,summarise,'Total'=sum(!is.na(City)),'< 20m'=sum(Points.Distance<20),'% <20m'=round((sum(Points.Distance<20))/(sum(!is.na(City)))*100,digits=2),'No. < 50m'=sum(Points.Distance<20),'% < 50m'=round((sum(Points.Distance<50))/(sum(!is.na(City)))*100,digits=2))
kable(results,caption="Table x - showing os_osm")
```

- As per literature there are big differences between areas e.g. in Bristol and Manchester over 80% of the Bus - Stops are included in the OSM data whereas in Cardiff it is only about 14%.
- There seems to be consistency in the number of railway stations in the OSM data at about 40% this could be down to the analysis method though as they were converted to point nodes {this might also be the issue with bus stations, may have to revisit the analysis of these!}
- Ignoring the anomalies of the stations and bus stops in Cardiff every other city/feature combination has over 50% coverage in the OSM data.
- Overall the figure of 64% is reasonable considering the anomalies that need looking at above.

*OSM data compared with Ground Survey*

```{r}
gsv_results=read.csv("../data/results/gsv_osm.csv", header=TRUE, stringsAsFactors = FALSE)

results=ddply(gsv_results,~City+Type,summarise,'Mean Distance'=mean(Points.Distance),'Total in OS'=sum(!is.na(City)),'No. under 20m'=sum(Points.Distance<20),'Percentage in OSM under 20m'=round((sum(Points.Distance<20))/(sum(!is.na(City)))*100,digits=2),'Mean distance under 20m'=mean(Points.Distance[Points.Distance<20]),'Standard Deviation distance under 20m'=sd(Points.Distance[Points.Distance<20]),'No. under 50m'=sum(Points.Distance<20),'Percentage in OSM under 50m'=round((sum(Points.Distance<50))/(sum(!is.na(City)))*100,digits=2),'Mean distance under 50m'=mean(Points.Distance[Points.Distance<50]),'Standard Deviation under 50m'=sd(Points.Distance[Points.Distance<50]))
kable(results,caption="Table x - showing gsv_osm")
```

## 4.3 Thematic Accuracy

*Inspection of tags of relevant OSM data features*
ADD TABLE AFTER DOING ANALYSIS IN R!!!


- {I need to add the sidewalk element attributes and possibly remove the entrance element}
- No crossings had a ‘length’ tag added to them but this is not in the OSM wiki.
- The majorities of crossings had a type added to them, though we see the differences in areas again with Manchester being just over 50% whereas Cardiff and Bristol have 80% and nearly 90%.
- The wheelchair accessible element needs removing as this is more for general accessibility
- The incline tag for steps is less than 10% for all areas this may be because it is hard to measure.
- The number of steps is very low with only Cardiff having over 2% coverage. This is something that could easily be added.
- The steps surface tag for all areas has about 5% coverage.

*Ground survey comparison of OSM tags*

ADD TABLE AFTER DOING IT!!!

## 4.4 Temporal Quality

*Inspection of the history of OSM elements*

ADD TABLE AFTER DOING ANALYSIS IN R!!!

- The mean version number is higher for the data relevant to a visually impaired person.
- There is also a higher percentage of elements with over 5 edits.
- These both point to potentially higher data quality
- The number of days since the last edit is more for the relevant in Bristol and Manchester but not for Cardiff again showing the differences in data over area.

## 4.5 Positional Accuracy ##

*Comparison of OSM data with Ordnance Survey data*
```{r}
os_results=read.csv("../data/results/os_osm.csv", header=TRUE, stringsAsFactors = FALSE)

results=ddply(os_results,~City+Type,summarise,'Mean Distance'=mean(Points.Distance),'Total in OS'=sum(!is.na(City)),'Mean distance under 20m'=mean(Points.Distance[Points.Distance<20]),'Standard Deviation distance under 20m'=sd(Points.Distance[Points.Distance<20]),'Mean distance under 50m'=mean(Points.Distance[Points.Distance<50]),'Standard Deviation under 50m'=sd(Points.Distance[Points.Distance<50]))

kable(results,caption="Table x - showing gsv_osm_distances")

```
